---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# EndpointR

<!-- badges: start -->

<!-- badges: end -->

EndpointR is a 'batteries included', open-source R package for connecting to various Application Programming Interfaces ([APIs](https://en.wikipedia.org/wiki/API){target="_blank"}) for Machine Learning model predictions.

> **TIP:** If you are an experienced programmer, or have experience with hitting APIs, consider going directly to [httr2](https://httr2.r-lib.org/reference/index.html)

# Installation

EndpointR will not be put on CRAN, so you can download and install the latest development version with the following code:

```{r, eval=FALSE}
library(EndpointR)
remotes::install_github("jpcompartir/EndpointR")
```

# Quick Starts

## Hugging Face - embeddings

Securely set your API key:

```{r, eval = FALSE}
set_api_key("HF_API_KEY")

endpoint_url <- httr2::secret_decrypt("kcZCsc92Ty7PuAk7_FdCcdOU_dlDpvWdDyfpwCWg-wW80eJxJPdQ68nz4V_0922SzSwM5_dYfpTOqsZ-GodUpLN4PQbwE73wlZkBCWIaIXc15g", "ENDPOINTR_KEY") # encrypted - for demo purposes

```

Embed a single text:

```{r, eval = FALSE}
embeddings <- hf_embed_text(
  text = "Convert this text to embeddings",
  endpoint_url = endpoint_url,
  key_name = "HF_API_KEY"
)
```

Embed a list of texts in batches:

```{r, eval = FALSE}
results <- hf_embed_batch(
  texts = c("Text one", "Text two", "Text three"),
  endpoint_url = endpoint_url,
  key_name = "HF_API_KEY",
  batch_size = 10,
  concurrent_requests = 5
)
```

Embed a data frame of texts:

```{r, eval = FALSE}
df_with_embeddings <- hf_embed_df(
  df = your_data,
  text_var = review_text,
  id_var = review_id,
  endpoint_url = endpoint_url,
  key_name = "HF_API_KEY"
)
```

## Hugging Face - Classification

Classify a single text:

```{r, eval = FALSE}
classification <- hf_classify_text(
  text = "This restaurant has terrible service",
  endpoint_url = your_classification_endpoint,
  key_name = "HF_API_KEY"
)
```

Classify a data frame:

```{r, eval = FALSE}

classified_df <- hf_classify_df(
  df = customer_reviews,
  text_var = review_text,
  id_var = review_id,
  endpoint_url = your_classification_endpoint,
  key_name = "HF_API_KEY",
  batch_size = 8,
  concurrent_requests = 3
)
```

Read the [Hugging Face Inference Vignette](vignettes/hugging_face_inference.Rmd) for more infromation on embedding and classifying using Dedicated Inference Endpoints and the Inference API from Hugging Face.

## OpenAI - Chat Completions API

Complete a single text:

```{r, eval = FALSE}
library(EndpointR)

set_api_key("OPENAI_API_KEY")

oai_complete_text(input = "Oh man, that's absolutely terrible.",
                  system_prompt = "Classify the sentiment of the following text: ")
```

Complete a Data Frame of texts:

```{r, eval = FALSE}
oai_complete_df(
  df = customer_reviews,
  text_var = review_text,
  id_var = review_id,
  system_prompt = "Classify the following review:",
  key_name = "OPENAI_API_KEY",
  concurrent_requests = 5 # send 5 rows of data simultaneously
)
```

Read the [LLM Providers Vignette](vignettes/llm_providers.Rmd), and the [Structured Outputs Vignette](vignettes/structured_outputs_json_schema.Rmd) for more information on common workflows with the OpenAI Chat Completions API [^1]

[^1]: Content pending implementation for Anthroic Messages API, Gemini API, and OpenAI Responses API

# API Key Security

-   Read the [httr2 vignette](https://httr2.r-lib.org/articles/wrapping-apis.html#basics){target="_blank"} on managing your API keys securely and encrypting them.

-   Read the [EndpointR API Keys](vignettes/api_keys.Rmd) vignette for information on which API keys you need for wach endpoint we support, and how to securely import those API keys into your .Renvironfile.
