---
title: "Using Hugging Face Inference Endpoints"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    number_sections: true
    fig_caption: true
    df_print: paged
    highlight: tango
    code_folding: show
    anchor_sections: true
vignette: >
  %\VignetteIndexEntry{Using Hugging Face Inference Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

```{r setup, eval = TRUE}
library(EndpointR)
library(dplyr)
library(tibble)
library(httr2)
library(purrr)
library(tidyr)
library(ggplot2)
```

## Who is this vignette for?

This vignette is for people who want to embed or classify data using pre-configured, dedicated Hugging Face Endpoints, assuming you have received your API key and the endpoint URL/details.

Experienced users looking for fine-grained control over their API requests will be better served by heading directly to the {[httr2](https://httr2.r-lib.org/index.html)} package.

## What is Hugging Face?

For the past few years [Hugging Face](https://huggingface.co/) has been the de-facto location for open-source AI. It is a place to save and share datasets & Machine Learning models alike. Importantly, we can train our own models and then upload them to the Hugging Face Hub for others to use.

## What are Dedicated Inference Endpoints?

Any model that is saved on the Hugging Face Hub can be connected to a Hugging Face Inference Endpoint allowing us to get predictions from that model using Hugging Face's [^1] hardware. Given that modern-day Machine Learning models tend to require a large amount of compute, and not everybody has powerful laptops to use the models locally, this is a democratising force and a useful tool to have at our disposal.

[^1]: or another inference provider's

### Common Use-cases

-   Text embeddings via Sentence Transformers
-   Sentiment analysis
-   Text classification
-   Image recognition
-   Question answering
-   Text generation

This vignette focuses on using Hugging Face endpoints for text embeddings.

## What is the Hugging Face Inference API?

The Hugging Face Inference API is a free service (with paid options) that provides access to thousands of models hosted on the Hub through simple HTTP requests. Unlike Dedicated Endpoints, we don't rent specific hardware - instead, we share computing resources with other users on Hugging Face's infrastructure.

This can lead to wait times which are unknowable in advance, for production use-cases it's generally recommended to use Dedicated Inference Endpoints. However, for casual usage and testing purposes, the Inference API should suffice.

> **TIP:** EndpointR's Hugging Face code should work with both Dedicated Inference Endpoints and the Inference API simply by changing the URL.

# Getting Started - Dedicated Inference Endpoints

## API Key Management

First read the [EndpointR API Keys](vignettes/api_keys.Rmd) vignette.

Assuming the endpoint has been set up, you'll need to retrieve your API key from Hugging Face (or from the team/department responsible for providing API keys) and store it securely in an environment variable.

`set_api_key` will ask you to input the value for your API key using {askpass}, this reduces [^2] the likelihood of your API key ending up in your code somewhere.

[^2]: but never reduces to zero, so always be vigilant!

```{r}
set_api_key("HF_TEST_API_KEY")
```

EndpointR functions that need API keys will have a `key_name` argument, you put the same name as you entered into `set_api_key` and EndpointR will retrieve it, without printing the contents to the console.

# Quick Start - Embeddings

EndpointR provides convenient functions for common embedding tasks without exposing the underlying request structure. EndpointR will handle retries, errors, concurrent requests, and batching of requests through configurable options.

## Dedicated Inference Endpoints - Embedding a Single text

> For this example, I have previously set up a package-level key for encrypting sensitive information, `ENDPOINTR_KEY`, this can be used with {httr2}'s `secret_encrypt` and `secret_decrypt` function to encrypt any information. For example, you can use these functions to encrypt your API keys themselves.

1.  Get your endpoint's URL
2.  Make sure the key_name points to the correct API key for your endpoint
3.  Get a single piece of text you want to embed
4.  Call `hf_embed_text()`

If the endpoint is not already active, it will take some time to initialise - the model's weights will be loaded onto a server. This process tends to take $\approx$ 20-30 seconds. Setting `max_retries = 5` will usually give the endpoint enough time to initialise, receive our request and return its response.

```{r}
endpoint_url <- httr2::secret_decrypt("kcZCsc92Ty7PuAk7_FdCcdOU_dlDpvWdDyfpwCWg-wW80eJxJPdQ68nz4V_0922SzSwM5_dYfpTOqsZ-GodUpLN4PQbwE73wlZkBCWIaIXc15g", "ENDPOINTR_KEY")

embeddings <- hf_embed_text(
  text = "This is a sample text to embed",
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY",
  max_retries = 5
)
```

The output is tidied into a tibble with 1 row and 384 columns. Each column is an embedding dimension, and the 'all-minilm-l6-v2' model used by this endpoint has 384 dimensions.

```{r}
embeddings
```

##Inference API - Embedding a Single Text

If we switch the URL that we send to our EndpointR functions, we can get predictions from the Hugging Face Inference API instead of Dedicated Inference Endpoints.

```{r}
url_inference_api <-"https://router.huggingface.co/hf-inference/models/sentence-transformers/all-mpnet-base-v2/pipeline/feature-extraction"

embeddings <- hf_embed_text(
  text = "This is a sample text to embed",
  endpoint_url = url_inference_api,
  key_name = "HF_TEST_API_KEY",
  max_retries = 5
)
```

## Working with Multiple Texts

When working with multiple texts, it would be possible to create a `hf_embed_text` function call for each text, and save each result to its own variable:

```{r}
text_1_embeddings <- hf_embed_text(text = "This is a sample text to embed", endpoint_url = endpoint_url, key_name = "HF_TEST_API_KEY", max_retries = 5)
text_2_embeddings <- hf_embed_text(text = "This is a sample text nunber 2 to embed", endpoint_url = endpoint_url, key_name = "HF_TEST_API_KEY", max_retries = 5)

```

And so on, for as many texts as we have. But this would get very boring, very quickly. Instead, it would be nice if we could embed multiple texts at the same time.

## Embedding a List of Texts {#embedding-a-list-of-texts}

Rather than send 10 requests and get 10 responses, we could batch our texts together so that each request has multiple texts. In the following example we'll send a list of 10 texts in 2 separate requests.

> **DISCLAIMER**: Claude was used to generate these sentences, instructed to write 10 sentences about embeddings, APIs, being British, and to add humour where it saw fit. The humour is 'mixed' at best.

```{r}
embedding_sentences <- c(
  "Text embedding models compress your rambling prose into compact vectors, much like how British commuters squeeze themselves into tube carriages during rush hour.",
  "Setting up an inference endpoint without proper documentation is akin to attempting to navigate London using a map of Birmingham.",
  "When your embedding model starts hallucinating, it's rather like watching Parliament during Question Time—entertaining but utterly unpredictable.",
  "Optimising your inference endpoint is essential unless you fancy your users growing old whilst waiting for a response, perhaps even ageing enough to collect their pension.",
  "The distance between word embeddings reveals semantic relationships, though sadly not the distance between what your client requested and what they actually wanted.",
  "Creating multilingual embeddings is a bit like attempting to order tea properly across Europe—technically possible but fraught with cultural misunderstandings.",
  "Batch processing through inference endpoints saves computing resources, much like how the British save conversation topics by discussing the weather exclusively.",
  "Token limits on embedding APIs are the digital equivalent of a queue at a British post office—inevitably, you'll reach the front just as they close for lunch.",
  "Fine-tuning embedding models on domain-specific corpora is rather like training a British child to apologise—it requires repetition, patience, and considerable examples.",
  "When your inference endpoint crashes under load, it maintains that quintessentially British trait of breaking down precisely when most inconvenient."
)
```

the `hf_embed_batch()` function handles a list of texts, and turns them into batches, sized according to what we enter into `batch_size =`. The default value is 8, we'll change it to 5 and send all of our texts in 2 requests.

```{r}
hf_embed_batch(
  embedding_sentences,
  endpoint_url,
  "HF_TEST_API_KEY",
  batch_size = 5,
)
```

> **TIP:** Remember, you could change `endpoint_url` to `url_inference_api` to send requests to the Hugging Face Inference API instead of a Dedicated Inference Endpoint.

The `hf_embed_batch` function will return our texts in a single data frame, with the text itself, two columns related to errors, and columns for each embedding dimension.

The `.error` and `.error_message` are introduced because we need to know whether errors occur at the individual post level. For example, if a batch of 10 requests has 2 errors, we want to see which individual post had errors, and keep the embeddings for the rest.

Batching is generally a good idea, and we'll see later that holding everything else equal, it does lead to a speed up over non-batched. However, if you try to set the batch_size too high, you are more likely to hit rate limits, or have your request rejected due to the payload being too large.

> **TIP**: Experiment to find the correct batch size for your data and the endpoint. Start small and work upwards in small incremeents.

## Embedding a Data Frame of Texts

As data professionals, we are more likely to work with data frames than any other data structure. We will usually have the texts we want to embed, their unique identifiers, and a selection of other columns to work with. EndpointR has a function - `hf_embed_df()` - built just for this use case.

For now, let's create some IDs and build a data frame from the sentences we defined earlier:

```{r}
embedding_ids <- 1:10

embedding_df <-
  tibble(
    id = embedding_ids,
    sentence = embedding_sentences
  )
```

The function requires an input to:

-   `text_var =` this should be the column in your data frame with the text data you need to embed.
-   `id_var =` this should be the column in your data frame which identifies each piece of text - this makes sure we can link our embeddings to the correct text.
-   `key_name` we've seen this before
-   `endpoint_url` we've seen this before

And then we have some optional arguments - these are optional because there is a default value for when we don't set them. We'll look at some of the other optional arguments in a later section. The key optional arguments for `hf_embed_df()` are:

-   `progress` lets EndpointR know whether to provide a progress bar, this is only really we're worth it if we're embedding a lot of data. For now we'll set it to `FALSE`
-   `max_retries` tells EndpointR how many times it should try to repeat the request before moving on to the next item.

> **NOTE**: We'll look at some additional arguments, such as batch_size and concurrent_requests later in [scaling performance](improving-performance)

```{r}
(
  result_df <- hf_embed_df(
  df = embedding_df,
  text_var = sentence,
  id_var = id,
  key_name = "HF_TEST_API_KEY",
  endpoint_url = endpoint_url,
  progress = FALSE,
  max_retries = 5)
)
```

We now have a data frame with our original ID column, the text column, the same '`.error` columns as in the batching case, and a number of columns with a 'V' followed by some digits. Each digit represents an 'embedding dimension', you should have as many 'VXX' columns as there are embedding dimensions for the model you are using.

You can check whether you had any errors a number of ways, I like to use `dplyr::count()` or `dplyr::filter()`, here's the count way:

```{r}
count(result_df, .error)
```

If you need to select only the columns that contain embeddings - for example if you want to feed these embeddings into a clustering, or dimensionality reduction model, you can select a range of columns using `dplyr::select()` - here I'll select V1:V384, which gives me all of the embeddings.

```{r}
result_df |> select(V1:V384)
```

If I had 768 embedding dimensions, I would input `V1:V768`, and similarly for 1024 dimensions, `V1:V1024`, you get the idea.

## Improving Performance

> This section will focus mainly on the `hf_embed_df()` function and the options we have for improving throughput of requests.

So far we've looked at how to send a single piece of text, a list or batch of texts, and how to iterate through a data frame full of texts, sending one request at a time. Whilst these are key workflows, once you're more comfortable with sending requests to APIs and receiving responses, you'll probably need to send a bunch of requests quickly, or in a more memory-efficient manner.

Before we get into how, let's go a bit further into why.

For each row of our data frame, we need to create a request which we will then perform. We then need to perform the request by sending our information over the internet to the endpoint. The endpoint accepts our request, performs any computation, and sends a successful response or an error. Each of these steps has an associated cost in time, and the cost for each step is not equal - some steps are more costly than others.

Creating each request is almost instant, but if we have a larger data frame - e.g. with 100,000 rows: creating 100,000 requests simultaneously will take $\approx$ 30s-90s. If we process them all at once, we'll have to wait the full amount of time before we send any requests and receive any responses. Clearly this is wasteful - if we had created the first request and sent it, we wouldn't be waiting 90s just to start sending requests.

Sending each request to the endpoint and waiting for it to process will usually take a lot longer than preparing the request. Therefore, if we can find ways to reduce the time we spend waiting here, we should be able to reduce the overall time it takes to prepare and send our requests significantly.

For the `hf_embed_df` function, our main options are to increase the number of requests we send at at once, and the number of texts we send within each request.

-   `concurrent_requests` lets EndpointR know how many requests to send at the same time. This should be a number between 1 and \~20-50 (in extreme cases you may be ok with 100, but most endpoints will not accept this)
-   `batch_size` lets EndpointR know whether to send requests with individual texts, or with a number of texts 'batched up' together.

## Concurrent Requests

Hugging Face' Inference Endpoints can handle multiple request arriving at the same time - and if configured they will 'autoscale' to provide higher throughput when there is a backlog of requests. As a rule, start with \~5 concurrent requests, and work up to \~20. If you start hitting rate limits go back down to \~10 and find the sweetspot. If endpoint is handling 20 requests then you could continue increasing gradually but it's not recommended unless you are sure the endpoint can handle it.

Here we send 5 concurrent requests - this will iterate through the data frame 5 rows at a time, and send new requests when responses are returned, until we run out of data to embed.

```{r}
hf_embed_df(
  df = embedding_df,
  text_var = sentence,
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY",
  id_var= id,
  concurrent_requests = 5,
  progress = TRUE
)
```

Let's benchmark performance (see appendix for code) - there is some overhead associated with generating parallel requests, so we'll need a bigger data frame to understand what type of speed up we can get.

```{r}
id_1000 <- 1:1000
sentences_1000 <- rep(embedding_df$sentence, 100)
embedding_df_1000 <- tibble(id = id_1000, sentence = sentences_1000)
```

Recording the results: we can see a \~40% reduction in processing time when going from 5-\> 10 concurrent requests and a \~15% reduction when going from 10 -\> 20 concurrent requests.

| concurrent_requests | processing_time_secs | success |
|--------------------:|---------------------:|:--------|
|                   5 |                 19.8 | TRUE    |
|                  10 |                 11.9 | TRUE    |
|                  20 |                 10.2 | TRUE    |

\*Exact times will fluctuate, so take these as approximates.

## Batch Requests

If our endpoint does not allow for multiple concurrent requests, or it's being overloaded by other users, we can still gain some efficiency by sending batches of the data frame in each request. Under the hood this looks quite similar to the [section](#embedding-a-list-of-texts) on embedding a list of texts in a batch.

> **TIP**: experiment with batch sizes to find the sweetspot - usually starting around 8-16, and capping out at \~64. You'll know when you've gone too high because you'll start seeing the retry bar, and/or your responses will contain errors.

```{r}
hf_embed_df(
  df = embedding_df_1000,
  text_var = sentence,
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY",
  id_var= id,
  concurrent_requests = 5,
  batch_size = 20,
  progress = TRUE
)
```

## Batching Concurrent Requests

For maximum speed up, we can also send batches of multiple concurrent requests. If our endpoint is able to handle them, 10 concurrent requests of `batch_size = 10` will be marginally faster than 10 concurrent requests of `batch_size = 5`.

### Benchmarking Batching & Concurrent Requests

In a separate session we did some benchmarking to understand the relationship between batch size, concurrent requests, and throughput, where throughput is the number of rows processed per second. We looked at combinations of:

-   **batch_size = c(1, 4, 8, 16, 32)**
-   **concurrent_requests = c(1, 5, 10, 15, 20)**

```{r, eval = TRUE}
data(batch_concurrent_benchmark, package = "EndpointR")

knitr::kable(batch_concurrent_benchmark |> mutate_if(is.numeric, ~ round(.x, 2)))
```

The results effectively speak for themselves. To embed $\approx$ 2,000 documents sending them 1 text and 1 request at a time, we get a throughput of $approx$ 2.16 texts per second, and it takes over 15 minutes! At the other end, 20 concurrent requests of batch size 8, and 20 concurrent requests of batch size 32 have a throughout of 195, which is close to 200x quicker. And we get our results back in 10 seconds!

> **NOTE**: If we needed to, we could re-run the benchmarking code multiple times but the general trend is very clear.

```{r, eval = TRUE}
batch_concurrent_benchmark |> 
  mutate(batch_size = factor(batch_size), concurrent_requests= factor(concurrent_requests)) |>
  ggplot(aes(x= batch_size, y = throughput, group = concurrent_requests)) +
  geom_point(aes(colour = concurrent_requests)) +
  geom_line(aes(colour = concurrent_requests)) +
  labs(y = "Throughput", x = "Batch Size", title = "Increasing `batch_size` and `concurrent_requests` increases throughput") +
  scale_colour_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

```{r, eval = TRUE}
batch_concurrent_benchmark |> 
    mutate(batch_size = factor(batch_size), 
           concurrent_requests = factor(concurrent_requests)) |>  
    ggplot(aes(x = batch_size, y = concurrent_requests, fill = throughput)) + 
    geom_tile() +
    geom_text(aes(label = round(throughput, 1)), colour = "white") +
    scale_fill_viridis_c(name = "Throughput") +
    theme_minimal() +
    labs(x = "Batch Size", y = "Concurrent Requests",
         title = "Throughput by Batch Size and Concurrent Requests") +
    theme(legend.position = "none")
```

Which parameter is more important for speed, `batch_size =` or `concurrent_requests =`? By eye it looks like concurrent_requests has the bigger effect, but it's clear that both parameters have a positive effect, and within the boundaries of our parameters, there is an approximately linear relationship between them and throughput.

Putting together a quick linear model confirms that everything else equal, increasing `concurrent_requests` increases throughput more than `batch_size`

```{r, eval = TRUE}
model <- lm(throughput ~ batch_size + concurrent_requests, data = batch_concurrent_benchmark)

broom::tidy(model) |>  
  mutate(across(c(estimate, std.error, statistic), ~round(.x, 2)))
```

> **QUESTION**: What might happen if we keep increasing `batch_size` and `concurrent_requests`? Will the relationship hold or not?

## Advanced Usage

The previous sections have explored a high-level API for single texts, lists of texts, and data frames. Sometimes you will need to do something the `hf_embed_text`, `hf_embed_batch` and `hf_embed_df` functions do not allow, and their 'helpful defaults' get in the way. I would generally recommend you go direct to {httr2}, but if you're intent on using EndpointR then material may be forthcoming on some of the lower-level functions, if you can't wait then head into the repo and explore the functions themselves!

# Quick Start - Classification

> **TIP**: this section of the vignette will assume familiarity with the section on embeddings, i.e. it's expected that you know what a Hugging Face Inference Endpoint is, and the various options common to this package like 'concurrent_requests', batch_size' etc.

Similarly to the embeddings section, EndpointR provides 3 separate high-level functions which try to help you classify text without exposing the underlying request structure.

-   `hf_classify_text()`
-   `hf_classify_batch()`
-   `hf_classify_df()`

With one text:

```{r}
endpoint_url <- httr2::secret_decrypt("c2nF_Tx_4V32AO0PZOohHvmKqXZ4Ysd0UfqL1loDrge5eIbLIF07ynGstlRtRJHUgeaMVROKOgnuiWs66VYqLM3SFIufcFv9vIDwOSQyTB0B5A", "ENDPOINTR_KEY")

text <- "Annual income twenty pounds, annual expenditure nineteen nineteen and six, result happiness. Annual income twenty pounds, annual expenditure twenty pounds ought and six, result misery."

hf_classify_text(
  text,
  endpoint_url,
  "HF_TEST_API_KEY"
)

```

With a list of texts:

```{r}
texts <- c(
  "It was the best of times, it was the worst of times, it was the age of wisdom, it was the age of foolishness.",
  "Whether I shall turn out to be the hero of my own life, or whether that station will be held by anybody else, these pages must show.",
  "There is nothing in the world so irresistibly contagious as laughter and good humour.",
  "The pain of parting is nothing to the joy of meeting again.",
  "No one is useless in this world who lightens the burdens of another."
)

hf_classify_batch(
  texts,
  endpoint_url,
  "HF_TEST_API_KEY",
  batch_size = 2
)
```

With a data frame of texts and IDs:

```{r}
text_df <- tibble(
  text = texts
) |> 
  mutate(id = row_number(), .before = 1)

hf_classify_df(
  text_df,
  text, 
  id,
  endpoint_url,
  "HF_TEST_API_KEY"
)
```

# Appendix

## Appendix - Embeddings

TODO: out of date post-refactoring \### Benchmarking Concurrent Requests

```{r, benchmark_concurrent}
run_benchmark <- function(num_concurrent, data, endpoint, key) {
  start_time <- Sys.time()
  res_df <- try(hf_embed_df(
                  df = data,
                  text_var = sentence, 
                  id_var= id,         
                  endpoint_url = endpoint,
                  key_name = key,
                  include_errors = FALSE, 
                  concurrent_requests = num_concurrent,
                  progress = FALSE 
                ), silent = TRUE) 
  processing_time <- Sys.time() - start_time
  success <- !inherits(res_df, "try-error") && nrow(res_df) == nrow(data)
  return(data.frame(
    concurrent_requests = num_concurrent,
    processing_time_secs = as.numeric(processing_time, units = "secs"),
    success = success
  ))
}

num_requests_vec <- c(5, 10, 20)
results_list <- lapply(num_requests_vec, function(n) {
  run_benchmark(
    num_concurrent = n,
    data = embedding_df_1000,
    endpoint = endpoint_url,
    key = "HF_TEST_API_KEY" 
  )
})

(summary_df <- do.call(rbind, results_list))
```

### Benchmarking Batch and Concurrent Requests

You won't be able to re-run this exactly as-is, as the data frame is not provided with the package. You could bring your own data if you wanted to do this.

```{r,benchmark_concurrent_batching}
trust <- readr::read_csv("~/data/trust/trust_slice_spam_classification.csv") |> 
  select(text) |>
  mutate(id = row_number()) |>
  filter(!is.na(text), text != "") # stop NAs and empty vals crashing anything

chunk_size <- 2000
total_chunks <- 20

# this chunking logic is actually rubbish.
trust_chunks <- trust |>
  mutate(chunk_id = ceiling(id / chunk_size)) |>
  group_split(chunk_id) |>
  head(total_chunks)

benchmark_params <- crossing(
  batch_size = c(1, 4, 8, 16, 32),
  concurrent_requests = c(1, 5, 10, 20)
) |>
  mutate(chunk_index = row_number() %% total_chunks + 1)

benchmark_results <- benchmark_params |>
  mutate(result = pmap(list(batch_size, concurrent_requests, chunk_index), function(bs, cr, ci) {
    current_chunk <- trust_chunks[[ci]]

    start_time <- Sys.time()

    res <- try(hf_embed_df(
      df = current_chunk,
      text_var = text,
      id_var = id,
      endpoint_url = endpoint_url,
      key_name = "HF_TEST_API_KEY",
      batch_size = bs,
      concurrent_requests = cr,
      progress = TRUE
    ), silent = TRUE)

    elapsed_time <- as.numeric(Sys.time() - start_time, units = "secs")
    success <- !inherits(res, "try-error")
    rows_processed <- if(success) nrow(current_chunk) else 0

    list(
      elapsed_time = elapsed_time,
      success = success,
      rows_processed = rows_processed,
      throughput = if(success) rows_processed / elapsed_time else 0
    )
  })) |>
  nest_wider(result)
```

## Appendix - Classification
