---
title: "Using Hugging Face Inference Endpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Hugging Face Inference Endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(EndpointR)
library(dplyr)
library(tibble)
library(httr2)
library(purrr)
```

# Who is this vignette for?

This vignette is for users needing to embed text using pre-configured company endpoints via the {EndpointR} package, assuming you have received your API key and the endpoint URL/details.

# What is Hugging Face?

For the past few years [Hugging Face](https://huggingface.co/) has been the de-facto location for open-source AI. It is a place to save and share datasets & Machine Learning models alike. Importantly, we can train our own models and then upload them to the Hugging Face Hub, so that other people can access them.

# What are Inference Endpoints?

Any model that is saved on Hugging Face can be connected to a Hugging Face Inference Endpoint allowing us to get predictions from that model using Hugging Face's [^1] hardware. Given that modern-day Machine Learning models tend to require a large amount of compute, and not everybody has powerful laptops to use the models locally, this is a democratising force and a useful tool to have at our disposal.

[^1]: or another inference provider's

## Common Use-cases

-   Text embeddings via Sentence Transformers
-   Sentiment analysis
-   Text classification
-   Image recognition
-   Question answering
-   Text generation

This vignette focuses on using Hugging Face endpoints for text embeddings.

# API Key Management

Assuming the endpoint has already been set up, you'll need to retrieve your API key from Hugging Face (or the team responsible for providing API keys) and store it securely in an environment variable.

`set_api_key` will ask you to input the value for your API key using {askpass}, this reduces [^2] the likelihood of your API key ending up in your code somewhere. EndpointR functions will then use a similar function to securely read the key, without printing it to your console.

[^2]: but never reduces to zero, so always be vigilant!

```{r}
set_api_key("HF_TEST_API_KEY")
```

# High-level API

The high-level API provides convenient functions for common embedding tasks without exposing the underlying request structure. Functions at this level will implement retries, error handling, and concurrent requests through configurable options.

## Embedding a Single Text

> For this example, I have previously set up a package-level key for encrypting sensitive information, `ENDPOINTR_KEY`, this can be used with {httr2}'s `secret_encrypt` and `secret_decrypt` function to encrypt any information. For example, you can use these functions to encrypt your API key itself.

1.  Get your endpoint's URL
2.  Make sure the key_name points to the correct API key for your endpoint
3.  Get a single piece of text you want to embed
4.  Call `hf_embed_text()`

If this is the first time we're hitting the endpoint, it will not be running. The endpoint initialises when we send a request with a valid API key. Behind the scenes the model's weights are loaded onto a server. This process tends to take \~20 seconds, so our first requests will always be delayed. `max_retries = 5` tells EndpointR to retry 5 times, with an increasing time increment between requests; this gives the endpoint enough time to initialise.

```{r}
endpoint_url <- httr2::secret_decrypt("kcZCsc92Ty7PuAk7_FdCcdOU_dlDpvWdDyfpwCWg-wW80eJxJPdQ68nz4V_0922SzSwM5_dYfpTOqsZ-GodUpLN4PQbwE73wlZkBCWIaIXc15g", "ENDPOINTR_KEY")

embeddings <- hf_embed_text(
  text = "This is a sample text to embed",
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY",
  max_retries = 5
)
```

The output is tidied into a tibble with 1 row and 384 columns, because the endpoint is using the 'all-minilm-l6-v2' model under the hood:

```{r}
embeddings
```

### Testing a response

If you are testing an API request with `hf_embed_text()`, you can pass arguments into `httr2::req_perform()`. For example, you can pass in a value for 'verbosity' to get more information on the request.

> How much information to print? This is a wrapper around req_verbose() that uses an integer to control verbosity: 0: no output 1: show headers 2: show headers and bodies 3: show headers, bodies, and curl status messages.

You can also pass in a value for 'path', which will save the response to a file.

## Embedding a Data Frame of Texts

When working with multiple texts, it would be possible to create a `hf_embed_text` function call for each text, and save each result to its own variable:

```{r}
text_1_embeddings <- hf_embed_text(text = "This is a sample text to embed", endpoint_url = endpoint_url, key_name = "HF_TEST_API_KEY", max_retries = 5)
text_2_embeddings <- hf_embed_text(text = "This is a sample text nunber 2 to embed", endpoint_url = endpoint_url, key_name = "HF_TEST_API_KEY", max_retries = 5)

```

And so on, for as many texts as we have. But this would very quickly get old. Instead we want to process multiple texts in one function call. As data professionals, we are often working with data frames, and given this is where our texts and their IDs will be stored, it makes sense to roll our data up into a data frame, and embed directly from this data frame, i.e. we want to put a data frame into the function and get a data frame back out. Whilst this is convenient, we run into some difficulties when working with large data frames. We'll look at how to work around these difficulties below when we talk about batching and sending multiple concurrent requests.

For now, let's create a data fame and embed some texts[^3].

[^3]: Claude was used to generate these sentences, instructed to write 10 sentences about embeddings, APIs, being British, and to add humour where it saw fit. The results are 'mixed' at best.

```{r}
embedding_ids <- 1:10
embedding_sentences <- c(
  "Text embedding models compress your rambling prose into compact vectors, much like how British commuters squeeze themselves into tube carriages during rush hour.",
  "Setting up an inference endpoint without proper documentation is akin to attempting to navigate London using a map of Birmingham.",
  "When your embedding model starts hallucinating, it's rather like watching Parliament during Question Time—entertaining but utterly unpredictable.",
  "Optimising your inference endpoint is essential unless you fancy your users growing old whilst waiting for a response, perhaps even ageing enough to collect their pension.",
  "The distance between word embeddings reveals semantic relationships, though sadly not the distance between what your client requested and what they actually wanted.",
  "Creating multilingual embeddings is a bit like attempting to order tea properly across Europe—technically possible but fraught with cultural misunderstandings.",
  "Batch processing through inference endpoints saves computing resources, much like how the British save conversation topics by discussing the weather exclusively.",
  "Token limits on embedding APIs are the digital equivalent of a queue at a British post office—inevitably, you'll reach the front just as they close for lunch.",
  "Fine-tuning embedding models on domain-specific corpora is rather like training a British child to apologise—it requires repetition, patience, and considerable examples.",
  "When your inference endpoint crashes under load, it maintains that quintessentially British trait of breaking down precisely when most inconvenient."
)

embedding_df <-
  tibble(
    id = embedding_ids,
    sentence = embedding_sentences
  )
```

Now we can the use the `hf_embed_df` to embed each of the 10 rows in our data frame. Due to the lower number of sentences, this will happen almost instantaneously, so long as the endpoint is already running. If the endpoint has shut down - if it's been a while between your requests, give it \~30 seconds and it will start back up.

When embedding our data frame, we must provide an input to:

-   `text_var =` this should be the column in your data frame with the text data you need to embed.
-   `id_var =` this should be the column in your data frame which identifies each piece of text - this makes sure we can link our embeddings to the correct text.
-   `key_name` we've seen this before
-   `endpoint_url` we've seen this before

There are some other optional arguments:

-   `concurrent_requests` lets EndpointR know whether to send multiple requests or not.
-   `progress` lets EndpointR know whether to provide a progress bar, this is only really we're worth it if we're embedding a lot of data. For now we'll set it to `FALSE`
-   `include_errors` should the rows where the API request failed by kept in the data, or not? We'll set this to `TRUE` so that we get a data frame with the same number of rows as the one we put in.
-   `max_retries` tells EndpointR how many times it should try to repeat the request before moving on to the next item.

```{r}
(
  result_df <- hf_embed_df(
  df = embedding_df,
  text_var = sentence,
  id_var = id,
  key_name = "HF_TEST_API_KEY",
  endpoint_url = endpoint_url,
  progress = FALSE,
  include_errors = TRUE,
  max_retries = 5)
)
```

We now have a data frame with our original ID column, the text column, and some new columns:

-   `.error` indicates whether there was an error or not
-   `.error_message` displays the content of the error if there was one

And then a number of columns with a 'V' followed by some digits. Each digit represents an 'embedding dimension', you should have as many 'VXX' columns as there are embedding dimensions for the model you are using.

You can check whether you had any erorrs a number of ways, I like to use `dplyr::count()` or `dplyr::filter()`, here's the count way:

```{r}
count(result_df, .error)
```

If you need to select only the columns that contain embeddings - for example if you want to feed these embeddings into a clustering, or dimensionality reduction model, you can select a range of columns using `dplyr::select()` - here I'll select V1:V384, which gives me all of the embeddings.

```{r}
result_df |> select(V1:V384)
```

If I had 768 embedding dimensions, I would input `V1:V768`, and similarly for 1024 dimensions, `V1:V1024`, you get the idea.

## Scaling Performance

So far we've looked at how to send a single piece of text, and how to iterate through a data frame full of texts, sending one request at a time. Whilst these are key workflows, once you're more comfortable with sending requests to APIs and receiving responses, you'll probably need to send a bunch of requests quickly, or in a more memory-efficient manner.

Before we get into how, let's go a bit further into why.

For each row of our data frame, we need to create a request which we will then perform. We then need to perform the request by sending our information over the internet to the endpoint. The endpoint accepts our request, performs any computation, and sends a successful response or an error. Each of these steps has a cost in time associated with it, and the cost for each step is not equal - some steps are more costly than others.

Creating each request is almost instant, but if we have a larger data frame - e.g. with 100,000 rows: creating 100,000 requests simultaneously will take $\approx$ 30s-90s. If we process them all at once, we'll have to wait the full amount of time before we send any requests and receive any responses. Clearly this is wasteful - if we had created the first request and sent it, we wouldn't be waiting 90s just to start sending requests.

Sending each request to the endpoint and waiting for it to process will usually take a lot longer than preparing the request. Therefore, if we can find ways to reduce the time we spend waiting here, we should be able to reduce the overall time it takes to prepare and send our requests significantly.

### Concurrent Requests

Hugging Face' Inference Endpoints can handle multiple request arriving at the same time - and if configured they will 'autoscale' to provide higher throughput when there is a backlog of requests. As a rule, start with \~5 concurrent requests, and work up to \~20. If you start hitting rate limits go back down to \~10 and find the sweetspot. If endpoint is handling 20 requests then you could continue increasing gradually but it's not recommended unless you are sure the endpoint can handle it.

Here we send 5 concurrent requests - this will iterate through the data frame 5 rows at a time, and send new requests when responses are returned, until we run out of rows.

```{r}
hf_embed_df(
  df = embedding_df,
  text_var = sentence,
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY",
  id_var= id,
  include_errors = TRUE,
  concurrent_requests = 5,
  progress = TRUE
)
```

Let's benchmark performance (see appendix for code) - there is some overhead associated with generating parallel requests, so we'll need a bigger data frame to understand what type of speed up we can get.

```{r}
id_1000 <- 1:1000
sentences_1000 <- rep(embedding_df$sentence, 100)
embedding_df_1000 <- tibble(id = id_1000, sentence = sentences_1000)
```

Recording the results: we can see a \~40% reduction in processing time when going from 5-\> 10 concurrent requests and a \~15% reduction when going from 10 -\> 20 concurrent requests.

| concurrent_requests | processing_time_secs | success |
|--------------------:|---------------------:|:--------|
|                   5 |                 19.8 | TRUE    |
|                  10 |                 11.9 | TRUE    |
|                  20 |                 10.2 | TRUE    |

\*Exact times will fluctuate, so take these as approximates.

### Batch Requests

Sending concurrent requests is one way to be more efficient - mainly through increasing throughput. Another way is to send more data inside each request, thus reducing total number of requests, and time spent sending requests to the endpoint.

To be concrete, this means sending multiple rows of data inside each request. If we have 1,000 rows of data, and we create batches of size 50, we'll have $\frac{1000}{50}= 20$ total requests to send, rather than the 1,000 requests we have without batching. If our endpoint is powered by GPU(s) or other hardware accelerators, it will take care of processing the data using parallel computation, increasing speed up.

> **Warning**: experiment with batch sizes to find the sweetspot - usually starting around 8-16, and capping out at \~64. You'll know when you've gone too high because you'll start seeing the retry bar, and/or your responses will contain errors.

TODO: come back to this when batching is fully fleshed out.

Going back to our list of sentences which we aved in the list - `embedding_sentences` above, we can create a single request with all 10 sentences batched in using `hf_build_request_batch()`

```{r}
batch_request <- hf_build_request_batch(
  embedding_sentences, endpoint_url, key_name = "HF_TEST_API_KEY")
```

Now if you inspect the body of the request, and look at data -\> inputs, you'll see the list of 10 sentences.

```{r}
batch_request$body$data
```

You can use `hf_perform_request()`to send the request. To the endpoint it will look like a single request, and it will return a single response. Each individual item in our requests `body$data$inputs` will be embedded separately.

```{r}
batch_result <- hf_perform_request(batch_request)
```

```{r}
result_df <- hf_embed_df(
  df = texts_df,
  text_var = text,
  id_var = id, 
  key_name = "HF_TEST_API_KEY",
  endpoint_url = endpoint_url,
  batch_size = 20,       # process 20 texts at a time
  progress = TRUE        # show progress bar
)
```

### Chunking Data Frames

When we have a lot of data - 100,000s or 1,000,000s of rows - we may struggle for memory if we try to create the requests all at once

### Writing to Files

Embeddings are returned to us as data frames of floating point numbers, where each column is an embedding dimension. The number of dimensions will usually fall in the range of \$\approx\$300-3000 (and some LLMs have internal represenations in the 10,000s of dimensions range!). That's a lot of data, and our computers only have so much RAM available. If we try to embed 100,000s - 1,000,000 rows at a time, there's a heightened chance of our R session crashing, meaning we lose the results. This wastes time, API credits, energy, and water.

It would be better to chunk our data up, create batches of requests, and write each batched response to a file. The file will persist if our R session crashes.

> At this stage it would normally be advisable to build up your own request directly with httr2 - you'll have full control over batching, concurrent requests, exponential back-off, time-outs and retries, error handling.

However, because it seems quite likely people will run into trouble, and there's a lot at stake if we're sending 100,000s-1,000,000s of requests, we've created a high-level function for chunking a data frame, batching requests, sending concurrent batches, and appending the batch to a file.

# Low-level API

The low-level API offers more control over the embedding process. This is useful for custom workflows, debugging, or when you need fine-grained control over requests.

## Request Preparation

Prepare requests without executing them, useful for inspection or custom processing:

```{r}
is_endpoint_active <- validate_hf_endpoint(endpoint_url, "HF_TEST_API_KEY")

req <- hf_embed_request_single(
  text = "Text for custom request",
  endpoint_url = endpoint_url,
  max_retries = 5,       # retry failed requests
  timeout = 30,          # longer timeout for large texts
  validate = FALSE,       # skip validation if already done
  key_name = "HF_TEST_API_KEY"
)

req # check it before you run it
```

## Manual Request Execution

If you've checked the request and everything looks ok, go ahead and execute it. We'll set tidy = FALSE first so that we can see what the raw request comes back like, then we'll tidy it up:

```{r}
response_raw <- hf_perform_single(req, tidy = FALSE)

resp_status <- httr2::resp_status(response_raw)
resp_headers <- httr2::resp_headers(response_raw)
```

We should now end up with our tidied data frame where each column is an embedding dimension.

```{r}
(embeddings <- tidy_embedding_response(response_raw))
```

## Batch Processing with Data Frames

Work directly with the component functions for custom batch workflows.

First we create a data frame of requests - this takes a data frame as an input, and returns a data frame but with some additional columns added.

```{r}
requests_df <- hf_build_request_df(
  df = texts_df,
  text_var = text,
  id_var = id,
  endpoint_url = endpoint_url,
  key_name = "HF_TEST_API_KEY"
)

str(requests_df$.request[[1]]) # check the structure of the first request
requests_df$.request[[1]]$body # check the data that's been added to the request
```

Then we batch our data frames, returning a list of `nrow(df)/batch_size` where each element in the list is its own data frame with `nrow(df) == batch_size`.

Still very much a WIP / TODO as to how this should work, too manual atm, probably.

```{r}
batch_dfs <- requests_df |>  batch_dataframe(5)
length(batch_dfs)
```

```{r}
responses_batch_dfs <- purrr::map(batch_dfs, ~ hf_perform_sequential(.x), .progress = TRUE)

purrr::map(responses_batch_dfs, ~ process_batch_responses(.x)) |> 
  bind_rows()
```

## Error Handling

The low-level API allows more detailed error inspection and recovery. We just wrap `req_perform` in `purrr::safely`. This changes the return object slightly, we now have a list with `$result` and `$error`

```{r}
result <- safely_perform_request(req)

if (!is.null(result$error)) {
  print(result$error)
} else {
  embeddings <- tidy_embedding_response(result$result)
}
```

# Appendix

## Benchmarking Concurrent Requests

```{r}
run_benchmark <- function(num_concurrent, data, endpoint, key) {
  start_time <- Sys.time()
  res_df <- try(hf_embed_df(
                  df = data,
                  text_var = sentence, 
                  id_var= id,         
                  endpoint_url = endpoint,
                  key_name = key,
                  include_errors = FALSE, 
                  concurrent_requests = num_concurrent,
                  progress = FALSE 
                ), silent = TRUE) 
  processing_time <- Sys.time() - start_time
  success <- !inherits(res_df, "try-error") && nrow(res_df) == nrow(data)
  return(data.frame(
    concurrent_requests = num_concurrent,
    processing_time_secs = as.numeric(processing_time, units = "secs"),
    success = success
  ))
}

num_requests_vec <- c(5, 10, 20)
results_list <- lapply(num_requests_vec, function(n) {
  run_benchmark(
    num_concurrent = n,
    data = embedding_df_1000,
    endpoint = endpoint_url,
    key = "HF_TEST_API_KEY" 
  )
})

(summary_df <- do.call(rbind, results_list))
```
