---
title: "Connecting to Major Model Providers"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    number_sections: true
    fig_caption: true
    df_print: paged
    highlight: tango
    code_folding: show
    anchor_sections: true
vignette: >
  %\VignetteIndexEntry{Connecting to Major Model Providers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(EndpointR)
library(httr2)
library(purrr)
library(dplyr)
```

This vignette details how we can use EndpointR to interact with services hosted by major model providers, such as Anthropic, Google, and OpenAI. The focus is on sending texts for a specific, one-off purpose - similar to how traditional Machine Learning endpoints work. Though, we will see that it's not always straightforward to get LLMs to do what we want them to do!

Before we get started, we need to make sure we have a couple of things in place:

-   First, get your API key and store it as "OPENAI_API_KEY" with `set_api_key()`
-   Second, figure out if you need the [Responses API](https://platform.openai.com/docs/api-reference/responses) or the [Chat Completions API](https://platform.openai.com/docs/api-reference/chat). Currently EndpointR only supports the Chat Completions API.
-   Third, choose your model - EndpointR is configured to select the smaller, cheaper model but you are free to choose specific models according to your needs.

> **Information:** if you are unsure whether you should prefer the Completions API or the Responses API, checkout [this webpage](https://platform.openai.com/docs/guides/responses-vs-chat-completions)

# OpenAI - Quick Start - Chat Completions API - Single Text

To get a completion for a single text we can use the `oai_complete_text()` function:

```{r}
set_api_key("OPENAI_API_KEY")

sentiment_system_prompt = "Analyse the sentiment of the given text."
text = "The weather has been absolutely fantastic this summer. I wish it could be like this every year, maybe I'll move to the South of France where they get 300 days of sunshine a year. Oh to dream."


oai_complete_text(
  text = text,
  system_prompt = sentiment_system_prompt,
  model = "gpt-4.1-nano"
)
```

Output: [1] "The sentiment of the text is positive. The speaker expresses happiness and admiration for the fantastic weather, along with a hopeful wish to experience similar conditions elsewhere."

## OpenAI - Quick Start - Chat Completions API - Data Frame of Texts

We can input a data frame directly into the `oai_complete_df()` function:

```{r}
review_df <- data.frame(
  id = 1:5,
  text = c(
    "Absolutely fantastic service! The staff were incredibly helpful and friendly.",
    "Terrible experience. Food was cold and the waiter was rude.",
    "Pretty good overall, but nothing special. Average food and service.",
    "Outstanding meal! Best restaurant I've been to in years. Highly recommend!",
    "Disappointed with the long wait times. Food was okay when it finally arrived."
  )
)

oai_complete_df(
  review_df,
  text_var = text,
  id_var = id,
  output_dir = NULL, # leave this to 'auto' to have your results written to a directory in your current working directory
  system_prompt = sentiment_system_prompt,
  concurrent_requests = 2, 
  chunk_size = 5
)
```

```{=html}
<pre>
ℹ Processing 5 texts in 1 chunk of up to 5 each
ℹ Performing 5 requests in parallel (with 2 concurrent requests)...
✔ Batch 1: 5 successful, 0 failed                                                                 
✔ Completed processing: 5 successful, 0 failed
# A tibble: 5 × 5                                                                                   
     id content                                                             .error .error_msg .chunk
  
<dbl> <chr>                                                               <lgl>  <lgl>       <dbl>
1     1 "The sentiment of the text is highly positive."                     FALSE  NA              1
2     2 "The sentiment of the text is negative."                            FALSE  NA              1
3     3 "The sentiment of the text is generally neutral with a slight lean… FALSE  NA              1
4     4 "The sentiment of the text is highly positive."                     FALSE  NA              1
5     5 "The sentiment of the text is negative."                            FALSE  NA              1
> 
</pre>
```

# OpenAI - Quick Start - Structured Outputs

An issue with our responses so far, is that they do not follow a specific format. This means we need to parse the results again after. However, we can use a schema to determine the format of the response:

```{r}
sentiment_schema <- create_json_schema(
  name = "simple_sentiment_schema",
  schema = schema_object(
    sentiment = schema_string(description = "Sentiment classification",
                              enum = c("positive", "negative", "neutral")),
    required = list("sentiment")
  )
)


structured_df <- oai_complete_df(
  review_df,
  text_var = text,
  id_var = id,
  schema = sentiment_schema,
  output_dir = NULL,
  system_prompt = sentiment_system_prompt,
  concurrent_requests = 2,
  chunk_size = 5
)

```

```{=html}
<pre>
 Processing 5 texts in 1 chunk of up to 5 each
ℹ Performing 5 requests in parallel (with 2 concurrent requests)...
✔ Batch 1: 5 successful, 0 failed                                                                 
✔ Completed processing: 5 successful, 0 failed
# A tibble: 5 × 5                                                                                   
     id content                        .error .error_msg .chunk
  
<dbl> <chr>                          <lgl>  <lgl>       <dbl>
1     1 "{\"sentiment\":\"positive\"}" FALSE  NA              1
2     2 "{\"sentiment\":\"negative\"}" FALSE  NA              1
3     3 "{\"sentiment\":\"neutral\"}"  FALSE  NA              1
4     4 "{\"sentiment\":\"positive\"}" FALSE  NA              1
5     5 "{\"sentiment\":\"negative\"}" FALSE  NA              1
> 
</pre>
```

Now we need to extract each sentiment value into a sentiment column, one per row:

```{r}
structured_df |>
  dplyr::mutate(content = purrr::map(content, ~safely_from_json(.x))) |> 
  tidyr::unnest_wider(content)
```

```{=html}
<pre>
# A tibble: 5 × 5
     id sentiment .error .error_msg .chunk
  <dbl> <chr>     <lgl>  <lgl>       <dbl>
1     1 positive  FALSE  NA              1
2     2 negative  FALSE  NA              1
3     3 neutral   FALSE  NA              1
4     4 positive  FALSE  NA              1
5     5 negative  FALSE  NA              1
</pre>
```

> **WARNING**: if you have a lot of data - i.e. 100,000s, or 1,000,000s of rows, you may want to split your data up and do this last step in chunks or via parallel processing to reduce peak memory usage.

# OpenAI - Under the Hood

We looked at `oai_complete_text()` and `oai_complete_df()` which are high-level convenience functions, but it's good to know what's happening under the hood.

## Chat Completions API

"The Chat Completions API endpoint will generate a model response from a list of messages comprising a conversation."

For most use-cases EndpointR covers, the conversation will be a single interaction between the user and the model. Usually the goal is to achieve some narrow task repeatedly.

> **TIP:** If you are looking for persistent, open-ended chats with LLMs, it's best to head to Claude, ChatGPT, Gemini - or your provider of choice.

So what type of tasks might we use the Completions API for?

-   Document classification, e.g. Sentiment Analysis
-   Translation
-   Structured data extraction

### Sentiment Analysis

Whilst we would generally not recommend using OpenAI's models for sentiment analysis, [^1] it is a task most people are familiar with.

[^1]: using such powerful models is inelegant - much smaller, less costly (time, energy, \$\$) models can do the job

We set up a very basic [^2] system prompt to tell the LLM what we want it to do with the text we send it:

[^2]: Generally prompts should be more detailed than this, we'll see why

```{r, sentiment_prompt}
system_prompt <- "Analyse the text's sentiment: "

text <- "Oh man, I'm getting to the end of my limit with this thing. WHY DOESN'T IT JUST WORK?!?"
```

Then we can create our request and inspect it with {httr2}'s `req_dry_run()` function - it's good hygiene to check your request is formatted as you expect it to be, particularly when using it for the first time. For example, if you intend to have a system prompt, is there a {"role": "system"} key:value pair inside the messages list? Is the value of "model": a valid model ID? Etc.

> **TIP**:You may need to familiarise yourself with the [OpenAI API Documentation](https://platform.openai.com/docs/api-reference/introduction) before proceeding.

```{r, sentiment_request}
sentiment_request <- oai_build_completions_request(
  text,
  system_prompt = system_prompt
) 

sentiment_request |>
  req_dry_run()
```

Your generated HTTP request should look something like this:

```{http}
POST /v1/chat/completions HTTP/1.1
accept: */*
accept-encoding: deflate, gzip
authorization: 
content-length: 247
content-type: application/json
host: api.openai.com
user-agent: EndpointR

{
  "model": "gpt-4.1-nano",
  "messages": [
    {
      "role": "system",
      "content": "Analyse the text's sentiment: "
    },
    {
      "role": "user",
      "content": "Oh man, I'm getting to the end of my limit with this thing. WHY DOESN'T IT JUST WORK?!?"
    }
  ],
  "temperature": 0,
  "max_tokens": 500
}
```

EndpointR (via {httr2}) handles the HTTP mechanics: authentication, request headers, and endpoint configuration. It then constructs the JSON payload with your specified model, system prompt, and user message, whilst setting sensible defaults for temperature [^3] and max_tokens [^4].

[^3]: lower values = less random output, max value is 2

[^4]: Maximum tokens of the *output* request

For demonstrative purposes, we ran the same prompt with the same data three times (see below)

```{r, sentiment_response}
sentiment_response <- sentiment_request |> 
  perform_request_or_return_error()
```

```{r, sentiment_response_text}
sentiment_response |>
  resp_check_status() |> 
  resp_body_json() |> 
  pluck("choices", 1, "message", "content")
```

First run:

> "The sentiment of the text is frustrated and exasperated."

Second run:

> "The sentiment of the text is quite negative. The speaker appears frustrated and exasperated, expressing dissatisfaction with the situation."

Third run:

> ""The sentiment of the text is quite negative. The speaker appears frustrated and exasperated, expressing dissatisfaction and impatience with the situation."

Whilst the responses are directionally/approximately accurate, the inconsistent output can be very unpleasant to work with. When scaled to hundreds or thousands of requests, this inconsistency can be extremely draining, and productivity reducing.

If we wanted the response to conform to the usual sentiment categories, we would need to write a custom parser to extract 'negative', 'neutral', or 'positive' from each output. Looking at the first output, it's clear the parser would need to be reasonably sophisticated. Or we could send another request, asking our model to please output only 'positive', 'negative', or 'neutral' [^5].

[^5]: or go to jail

Clearly this is more work than we should be willing to do. We'll look at techniques for how to deal with this systematically, and achieve predictable outputs in the [Structured Outputs](#openai-structured-outputs) section.

What does the model do when we hand it a text which is difficult for traditional, three-category [^6], document-level sentiment analysis?

[^6]: positive, negative, neutral

```{r, ambiguous_text}
ambiguous_text <- "The interface is brilliant but the performance is absolutely dreadful"

ambiguous_sentiment <- oai_build_completions_request(
  ambiguous_text,
  system_prompt = system_prompt
) |> 
  perform_request_or_return_error() |> 
  resp_body_json() |> 
  pluck("choices", 1, "message", "content")
```

[1] "The sentiment of the text is mixed, with positive feelings expressed about the interface ("brilliant") and negative feelings about the performance ("absolutely dreadful")."

The model is smart enough to recognise that the sentiment does not fit neatly into 'positive', 'negative', 'neutral' but once again the output is not formatted in a nice way for downstream use.

### Multiple Texts

Individual requests will be useful for some applications but EndpointR was built to handle more, whilst taking care of implementation details like concurrent requests, retries, and failing gracefully.

Let's experiment with a better system prompt and a list of texts:

```{r, updated_sentiment_prompt}
updated_sentiment_prompt <- "Classify the text into sentiment categories.
The accepted categories are 'positive', 'negative', 'neutral', and 'mixed'.
A 'mixed' text contains elements of positive and negative.
"
```

```{r, classical_texts}
classical_texts <- c(
 "It is a truth universally acknowledged, that a single man in possession of a good fortune, must be in want of a wife.",
 "All happy families are alike; each unhappy family is unhappy in its own way.",
 "It was the best of times, it was the worst of times.",
 "Call me Ishmael.",
 "The sun shone, having no alternative, on the nothing new.",
 "All animals are equal, but some animals are more equal than others.",
 "So we beat on, boats against the current, borne back ceaselessly into the past.",
 "The heart was made to be broken.",
 "Tomorrow, and tomorrow, and tomorrow, creeps in this petty pace from day to day.",
 "I have always depended on the kindness of strangers."
)
```

```{r, classical_requests}
classical_requests <- oai_build_completions_request_list(
  classical_texts,
  system_prompt = updated_sentiment_prompt
)
```

When running this in testing, it took 8.5 seconds for the 10 requests if we send one at a time.

```{r, seq_test}
start_seq <- Sys.time()
classical_responses  <- classical_requests |> 
  perform_requests_with_strategy()
end_seq <- Sys.time() - start_seq
```

And it took 1.8 seconds we send 10 requests in parallel - a \~4.7x speed up, showing there is some overhead cost in sending requests in parallel - i.e. we did not see a 10x speed increase for 10x requests.

```{r, par_test}
start_par <- Sys.time()
classical_responses  <- classical_requests |> 
  perform_requests_with_strategy(concurrent_requests = 10)
end_par <- Sys.time() - start_par
```

Now when we extract the content of the response, we can see [^7] that each response is a classification belonging to our classes - making our prompt slightly better helped us get to a better final result. However, if we were to repeat this over many texts - hundreds/thousands, it's unlikely every single response would conform to our categories. Without further instruction, the models have a tendency to slightly change the output at unpredictable intervals.

[^7]: at least when I ran this in testing

```{r}
classical_responses |> 
  map(~ resp_body_json(.x) |> 
        pluck("choices", 1, "message", "content") |> 
        as_tibble()
  ) |> 
  list_rbind() |> 
  mutate(text = classical_texts, .before = 1)
```

### Data Frame of Texts

```{r}
df_classical_texts <- tibble(
  id = 1:10,
  text = classical_texts
)
```

We have a function `oai_complete_df()` which takes a data frame, an id variable, and a text variable as mandatory inputs, and returns a data frame with columns: `id_var`, `content`, `.error`, `.error_msg`, `.chunk`.

```{r}
oai_complete_df(df_classical_texts, 
                text_var = text, 
                id_var = id,
                concurrent_requests = 5,
                output_dir = NULL # set this to write to a temporary directory, useful for documentation and testing.
                )
```

The `oai_complete_df()` function has a few knobs that you can turn to improve throughput, reduce memory usage, and increase the likelihood of a successful response. By default it will split your data frame into `chunk_size =` chunks. Each chunk is then processed in order, and a file is created to store the results of the previous batches. At the end of all chunks the results are retrieved from the file.

This should help to reduce the likelihood you:

1.  Lose the API responses of long-running calculations which fail part-way through
2.  Run out of memory as the responses list grows with the size of your data frame.

The function reports on progress as it happens, letting you know which chunk is being processed, e.g. 1/5, and a progress bar for the requests within that chunk. The function provides a summary of failures and successes directly in your console after each chunk, and at the end of all chunks.

## OpenAI Structured Outputs {#openai-structured-outputs}

The textual responses we get from LLM providers are difficult to deal with programmatically, because they make no guarantees on the form of the response. For example, if we ask the LLM to classify documents, sometimes it will give just the classification - 'positive', sometimes it will add some pre-amble 'the document is positive', and sometimes it will do something else entirely.

For detailed information on creating JSON schemas for structured outputs, see `vignette("structured_outputs_json_schema")`.

# Anthropic - Quick Start - Messages API - Single Text

Before getting started with Anthropic:

-   First, get your API key from the [Anthropic Console](https://console.anthropic.com/) and store it as "ANTHROPIC_API_KEY" with `set_api_key()`
-   Second, choose your model - EndpointR defaults to "claude-haiku-4-5" for cost efficiency, but you can use any Claude model
-   Third, note that Anthropic's temperature range is 0-1 (unlike OpenAI's 0-2)

To get a completion for a single text we can use the `ant_complete_text()` function:

```{r}
set_api_key("ANTHROPIC_API_KEY")

sentiment_system_prompt = "Analyse the sentiment of the given text."
text = "The weather has been absolutely fantastic this summer. I wish it could be like this every year, maybe I'll move to the South of France where they get 300 days of sunshine a year. Oh to dream."

ant_complete_text(
  text = text,
  system_prompt = sentiment_system_prompt,
  model = "claude-haiku-4-5"
)
```

Output: [1] "The sentiment of the text is positive and wistful. The author expresses happiness about the fantastic summer weather and daydreams about experiencing similar conditions year-round in the South of France."

## Anthropic - Quick Start - Messages API - Data Frame of Texts

We can input a data frame directly into the `ant_complete_df()` function:

```{r}
review_df <- data.frame(
  id = 1:5,
  text = c(
    "Absolutely fantastic service! The staff were incredibly helpful and friendly.",
    "Terrible experience. Food was cold and the waiter was rude.",
    "Pretty good overall, but nothing special. Average food and service.",
    "Outstanding meal! Best restaurant I've been to in years. Highly recommend!",
    "Disappointed with the long wait times. Food was okay when it finally arrived."
  )
)

ant_complete_df(
  review_df,
  text_var = text,
  id_var = id,
  output_dir = NULL, # leave this to 'auto' to have results written to a timestamped directory
  system_prompt = sentiment_system_prompt,
  concurrent_requests = 5,
  chunk_size = 5
)
```

```{=html}
<pre>
ℹ Processing 5 texts in 1 chunk of up to 5 each
ℹ Results will be saved as parquet files in [directory]
Processing chunk 1/1 (5 texts)
✔ Chunk 1: 5 successful, 0 failed
ℹ Processing completed, there were 5 successes and 0 failures.
# A tibble: 5 × 6
     id content                                                .error .error_msg .status .chunk
  <dbl> <chr>                                                  <lgl>  <chr>        <int>  <int>
1     1 "The sentiment is highly positive."                    FALSE  NA              NA      1
2     2 "The sentiment is negative."                           FALSE  NA              NA      1
3     3 "The sentiment is neutral with a slight positive lean." FALSE  NA              NA      1
4     4 "The sentiment is extremely positive."                 FALSE  NA              NA      1
5     5 "The sentiment is negative."                           FALSE  NA              NA      1
>
</pre>
```

# Anthropic - Quick Start - Structured Outputs

Anthropic supports [structured outputs](https://platform.claude.com/docs/en/build-with-claude/structured-outputs) (currently in beta) for the Claude 4.5 series models: `claude-haiku-4-5`, `claude-sonnet-4-5`, and `claude-opus-4-5`. Like with OpenAI, we can use a schema to ensure predictable, parseable responses:

> **NOTE**: Structured outputs are in beta and may change faster than we can track in this documentation. Check the [official documentation](https://platform.claude.com/docs/en/build-with-claude/structured-outputs) for the latest requirements.

```{r}
sentiment_schema <- create_json_schema(
  name = "simple_sentiment_schema",
  schema = schema_object(
    sentiment = schema_string(description = "Sentiment classification",
                              enum = c("positive", "negative", "neutral")),
    required = c("sentiment")
  )
)

structured_df <- ant_complete_df(
  review_df,
  text_var = text,
  id_var = id,
  schema = sentiment_schema,
  output_dir = NULL,
  system_prompt = sentiment_system_prompt,
  concurrent_requests = 5,
  chunk_size = 5
)
```

```{=html}
<pre>
ℹ Processing 5 texts in 1 chunk of up to 5 each
ℹ Results will be saved as parquet files in [directory]
Processing chunk 1/1 (5 texts)
✔ Chunk 1: 5 successful, 0 failed
ℹ Processing completed, there were 5 successes and 0 failures.
# A tibble: 5 × 6
     id content                        .error .error_msg .status .chunk
  <dbl> <chr>                          <lgl>  <chr>        <int>  <int>
1     1 "{\"sentiment\":\"positive\"}" FALSE  NA              NA      1
2     2 "{\"sentiment\":\"negative\"}" FALSE  NA              NA      1
3     3 "{\"sentiment\":\"neutral\"}"  FALSE  NA              NA      1
4     4 "{\"sentiment\":\"positive\"}" FALSE  NA              NA      1
5     5 "{\"sentiment\":\"negative\"}" FALSE  NA              NA      1
>
</pre>
```

Now we extract the sentiment values:

```{r}
structured_df |>
  dplyr::mutate(content = purrr::map(content, ~safely_from_json(.x))) |>
  tidyr::unnest_wider(content)
```

```{=html}
<pre>
# A tibble: 5 × 6
     id sentiment .error .error_msg .status .chunk
  <dbl> <chr>     <lgl>  <chr>        <int>  <int>
1     1 positive  FALSE  NA              NA      1
2     2 negative  FALSE  NA              NA      1
3     3 neutral   FALSE  NA              NA      1
4     4 positive  FALSE  NA              NA      1
5     5 negative  FALSE  NA              NA      1
</pre>
```

# Anthropic - Under the Hood

We looked at `ant_complete_text()` and `ant_complete_df()` which are high-level convenience functions, but let's understand what's happening under the hood.

## Messages API

Anthropic's Messages API is similar to OpenAI's Chat Completions API - it generates model responses from a conversation of messages. For EndpointR use cases, this is typically a single user-model interaction for a specific task.

### Building Requests

Let's create a request and inspect it:

```{r}
sentiment_text <- "This is absolutely brilliant! I couldn't be happier with the results."

sentiment_request <- ant_build_messages_request(
  input = sentiment_text,
  system_prompt = sentiment_system_prompt,
  model = "claude-haiku-4-5"
)

sentiment_request |>
  req_dry_run()
```

Your generated HTTP request should look something like this:

```{http}
POST /v1/messages HTTP/1.1
Host: api.anthropic.com
User-Agent: EndpointR
Content-Type: application/json
x-api-key: [YOUR_API_KEY]
anthropic-version: 2023-06-01

{
  "model": "claude-haiku-4-5",
  "messages": [
    {
      "role": "user",
      "content": "This is absolutely brilliant! I couldn't be happier with the results."
    }
  ],
  "max_tokens": 500,
  "temperature": 0,
  "system": "Analyse the sentiment of the given text."
}
```

Key differences from OpenAI: - **Authentication**: Anthropic uses `x-api-key` header (not `Authorization: Bearer`) - **System prompts**: Passed as a separate `system` parameter (not in messages array) - **Temperature range**: 0-1 for Anthropic (vs 0-2 for OpenAI) - **Required parameter**: `max_tokens` is required for Anthropic requests

### Structured Outputs with Anthropic

Anthropic's structured outputs use the `output_format` parameter with JSON schemas:

```{r}
structured_request <- ant_build_messages_request(
  input = sentiment_text,
  system_prompt = sentiment_system_prompt,
  schema = sentiment_schema,
  model = "claude-haiku-4-5"
)
```

When a schema is provided, the request automatically includes: - The `output_format` parameter with your schema - The `anthropic-beta` header for structured outputs support

The schema is validated and formatted automatically by EndpointR's `create_json_schema()` helper.

### Processing Multiple Texts

Like with OpenAI, we can process multiple texts efficiently with concurrent requests:

```{r}
classical_texts <- c(
  "It is a truth universally acknowledged, that a single man in possession of a good fortune, must be in want of a wife.",
  "All happy families are alike; each unhappy family is unhappy in its own way.",
  "It was the best of times, it was the worst of times."
)

# Build requests for each text
requests <- purrr::map(
  classical_texts,
  ~ant_build_messages_request(
    input = .x,
    system_prompt = "Classify the sentiment as positive, negative, neutral, or mixed."
  )
)

# Send all requests concurrently
responses <- perform_requests_with_strategy(
  requests,
  concurrent_requests = 3
)

# Extract content from responses
responses |>
  purrr::map_chr(~.extract_ant_message_content(.x))
```

### Memory-Efficient Batch Processing

The `ant_complete_chunks()` and `ant_complete_df()` functions handle large-scale processing by:

1.  Dividing input into chunks (default 5000 texts per chunk)
2.  Processing each chunk with concurrent requests
3.  Writing results to parquet files immediately
4.  Tracking metadata including the full schema structure

This approach minimises memory usage and prevents data loss if processing is interrupted.

**Key advantage**: Anthropic's metadata includes the complete schema definition (not just a boolean flag), making it easy to reproduce results and debug issues.

# Google

TBC
