---
title: "Synchronous vs Asynchronous APIs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synchronous vs Asynchronous APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EndpointR)
```

# Introduction

Most of EndpointR's integrations are with synchronous APIs such as [Completions](https://platform.openai.com/docs/api-reference/completions) by OpenAI, Hugging Face's [Inference Endpoints](https://huggingface.co/docs/inference-endpoints/en/index), and Messages by [Anthropic](https://platform.claude.com/docs/en/api/messages). When using these APIs, we send a request, wait a second or two and receive a response.

However, most Generative AI providers also offer lower-cost, asynchronous APIs. The providers usually offer a guarantee of the results within a time frame, and an estimate of the average time to return the results. For example, they may guarantee results within 24 hours, but expect them within 1-3 hours.

# When to choose Synchronous vs Asynchronous

> For a more comprehensive treatment, and motivating examples [OpenAI's offficial documentation/guide](https://platform.openai.com/docs/guides/batch) is a good place to start.

As consumers, the decision represents a trade-off between time and money. If we are serving Generative AI to other consumers, e.g. in an application, we will usually favour a Synchronous API because users expect instant results. Alternatively, if we are running analyses over large datasets, or repeated batch-inference, we can usually afford to wait longer for the results so we may favour the Asynchronous APIs.

Synchronous APIs are also very useful when we are still in the experimental step of an analysis and need quick feedback- i.e. when we're testing our prompts, and developing our [schemas](https://json-schema.org/) for [Structured Outputs](https://platform.openai.com/docs/guides/structured-outputs), we don't want to wait 24 hours to find out we made an error in our schema. Instead, we want to send a request and receive a response within seconds. That way we can iteratively fix/develop our schemas/prompts and get to a better outcome quicker.

At the time of writing, [OpenAI's batch API](https://platform.openai.com/docs/guides/batch) offers a 50% discount on the Completions API, as well as higher rate limits:

<blockquote>

Learn how to use OpenAI's Batch API to send asynchronous groups of requests with 50% lower costs, a separate pool of significantly higher rate limits, and a clear 24-hour turnaround time. The service is ideal for processing jobs that don't require immediate responses.

Batch processing jobs are often helpful in use cases like:

1.  Running evaluations
2.  Classifying large datasets
3.  Embedding content repositories

</blockquote>

# EndpointR Implementation of OpenAI Batch API

Due to inherent differences between Synchronous and Asynchronous APIs, the EndpointR implementation of the OpenAI Batch API will feel more like submitting jobs to a cluster/serve than automagically working with an entire data frame as in `oai_complete_df()` and `oai_embed_df()`.

\`oai_batch_create

`oai_prepare_batch()` `oai_prepare_batches()`

`oai_create_batch()` `oai_create_batches()`

`oai_check_batch_jobs()`
