---
title: "Synchronous vs Asynchronous/Batch APIs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synchronous vs Asynchronous APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}  
library(EndpointR)
```

# Introduction

Most of EndpointR's integrations are with synchronous APIs such as [Completions](https://platform.openai.com/docs/api-reference/completions) by OpenAI, Hugging Face's [Inference Endpoints](https://huggingface.co/docs/inference-endpoints/en/index), and Messages by [Anthropic](https://platform.claude.com/docs/en/api/messages). When using these APIs, we send a HTTP request, wait a second or two and receive a response.

However, data scientists often need to process an entire data frame, resulting in thousands or millions of HTTP requests. This is inefficient because:

1.  Cost - Providers don't offer discounts for these requests
2.  Session Blocking - Our coding consoles get blocked for hours at a time
3.  Rate Limits - Providers enforce stricter rate limits on these APIs

A solution to these problems is to use providers' 'Batch APIs' which offer asynchronous results. These often come with a 50% discount and higher rate limits, with a guarantee of results within a time frame, e.g. 24 hours.

> **TIP**: It's worth noting that the results are often ready much faster, consider checking in 1-2 hours after triggering the batch.

# Quickstart

The OpenAI Batch API workflow follows three stages: **prepare**, **submit**, and **retrieve**. Below are complete examples for embeddings and completions.

## Batch Embeddings

```{r batch-embeddings, eval = FALSE}
# 1. Prepare your data
df <- data.frame(
 id = c("doc_1", "doc_2", "doc_3"),
 text = c(
   "The quick brown fox jumps over the lazy dog",
   "Machine learning is transforming data science",
   "R is a powerful language for statistical computing"
 )
)

# 2. Prepare requests for the Batch API
jsonl_content <- oai_batch_prepare_embeddings(
 df,
 text_var = text,
 id_var = id,
 model = "text-embedding-3-small",
 dimensions = 256
)

# 3. Upload to the Files API
file_info <- oai_batch_file_upload(jsonl_content)
file_info$id
#> "file-abc123..."

# 4. Trigger the batch job
batch_job <- oai_batch_start(
 file_id = file_info$id,
 endpoint = "/v1/embeddings"
)
batch_job$id
#> "batch-xyz789..."

# 5. Check status (repeat until completed)
status <- oai_batch_status(batch_job$id)
status$status
#> "in_progress" ... later ... "completed"

# 6. Download and parse results
content <- oai_file_content(status$output_file_id)
embeddings_df <- oai_batch_parse_embeddings(content)

# Result: tidy data frame with id and embedding dimensions (V1, V2, ..., V256)
embeddings_df
#> # A tibble
#>   custom_id .error .error_msg    V1      V2      V3 ...
#>   <chr>     <lgl>  <chr>      <dbl>   <dbl>   <dbl> ...
#> 1 doc_1     FALSE  NA         0.023  -0.041   0.018 ...
#> 2 doc_2     FALSE  NA        -0.015   0.032   0.044 ...
#> 3 doc_3     FALSE  NA         0.008  -0.027   0.031 ...
```

## Batch Completions

```{r batch-completions, eval = FALSE}
# 1. Prepare your data
df <- data.frame(
 id = c("q1", "q2", "q3"),
 prompt = c(
   "What is the capital of France?",
   "Explain photosynthesis in one sentence.",
   "What is 2 + 2?"
 )
)

# 2. Prepare requests
jsonl_content <- oai_batch_prepare_completions(
 df,
 text_var = prompt,
 id_var = id,
 model = "gpt-4o-mini",
 system_prompt = "You are a helpful assistant. Be concise.",
 temperature = 0,
 max_tokens = 100
)

# 3. Upload and trigger batch job
file_info <- oai_batch_file_upload(jsonl_content)
batch_job <- oai_batch_start(
 file_id = file_info$id,
 endpoint = "/v1/chat/completions"
)

# 4. Check status and retrieve results
status <- oai_batch_status(batch_job$id)
# ... wait for status$status == "completed" ...

content <- oai_file_content(status$output_file_id)
completions_df <- oai_batch_parse_completions(content)

completions_df
#> # A tibble
#>   custom_id content                                      .error .error_msg
#>   <chr>     <chr>                                        <lgl>  <chr>
#> 1 q1        The capital of France is Paris.              FALSE  NA
#> 2 q2        Photosynthesis converts sunlight into energy FALSE  NA
#> 3 q3        2 + 2 equals 4.                              FALSE  NA
```

## Batch Completions with Structured Output

For classification tasks or when you need structured data back, combine the Batch API with JSON schemas:

```{r batch-completions-schema, eval = FALSE}
# 1. Define a schema for sentiment classification
sentiment_schema <- create_json_schema(
 name = "sentiment_analysis",
 schema_object(
   sentiment = schema_enum(
     c("positive", "negative", "neutral"),
     description = "The sentiment of the text"
   ),
   confidence = schema_number(
     description = "Confidence score between 0 and 1"
   )
 )
)

# 2. Prepare data
df <- data.frame(
 id = c("review_1", "review_2", "review_3"),
 text = c(
   "This product is absolutely fantastic! Best purchase ever.",
   "Terrible quality, broke after one day. Complete waste of money.",
   "It's okay, nothing special but does the job."
 )
)

# 3. Prepare requests with schema
jsonl_content <- oai_batch_prepare_completions(
  df,
  text_var = text,
  id_var = id,
  model = "gpt-4o-mini",
  system_prompt = "Analyse the sentiment of the following text.",
  schema = sentiment_schema,
  temperature = 0
)

# 4. Upload and trigger batch job
file_info <- oai_batch_file_upload(jsonl_content)
batch_job <- oai_batch_start(
 file_id = file_info$id,
 endpoint = "/v1/chat/completions"
)

# 5. Retrieve and parse results
status <- oai_batch_status(batch_job$id)
content <- oai_file_content(status$output_file_id)
results_df <- oai_batch_parse_completions(content)

# The content column contains JSON that can be parsed
results_df$content
#> [1] "{\"sentiment\":\"positive\",\"confidence\":0.95}"
#> [2] "{\"sentiment\":\"negative\",\"confidence\":0.92}"
#> [3] "{\"sentiment\":\"neutral\",\"confidence\":0.78}"

# Parse the JSON content into columns
results_df |>
 dplyr::mutate(
   parsed = purrr::map(content, jsonlite::fromJSON)
 ) |>
 tidyr::unnest_wider(parsed)
#> # A tibble
#>   custom_id sentiment confidence .error .error_msg
#>   <chr>     <chr>          <dbl> <lgl>  <chr>
#> 1 review_1  positive        0.95 FALSE  NA
#> 2 review_2  negative        0.92 FALSE  NA
#> 3 review_3  neutral         0.78 FALSE  NA
```

> **Limits**: Each batch file can contain up to 50,000 requests or 200MB, whichever is reached first. For larger datasets, split into multiple batches.

# When to choose Synchronous vs Asynchronous

> For a more comprehensive treatment, and motivating examples [OpenAI's official documentation/guide](https://platform.openai.com/docs/guides/batch) is a good place to start.

|   | Synchronous | Asynchronous (Batch) |
|----|----|----|
| Cost | Full price per token | \~50% Discount per token |
| Latency | Real-time | Up to 24 hours |
| Use Case | Experimentation, Prompt testing,  Schema development, User-facing applications, | Recurrent workflows (evals etc.), embedding large datasets, classifying large datasets |
| Data Size | Up to \~10,000 | \~10,000+ |

> **Recommendation**: Use the Synchronous API when you need immediate feedback e.g. prompt or schema development, and for small datasets where cost savings are irrelevant. Once everything is figured out, move to the Batch API to save on cost.

# Cleaning Up

Once the batch job has been completed, the associated files will live on the OpenAI API, inside the Files API. Your OpenAI account will be charged for storage, so it's best to download the results and save in your org's own cloud storage.

```{r, eval = FALSE}
oai_file_delete(file_info$id) # delete the input file 

oai_file_delete(status$output_file_id) # delete the output file
oai_file_delete(status$error_file_id) # delete the error file
```


> **NOTE**: At the time of writing, OpenAI save information in both the Batch API and the Files API, you need to delete your input, output, error files from the *Files API*, you cannot delete from the Batch API

# Technical Details

## Batch Limits

The OpenAI Batch API enforces specific limits per batch file. If your data exceeds these, you must split it into multiple batch jobs.

- Max Requests per Batch: 50,000
- Max File Size: 200 MB
  
  > **Warning**: When using Structured Outputs, the JSON schema is repeated for every single request in the batch file. For complex schemas, you may hit the 200 MB file size limit well before you reach the 50,000 row limit.

## Underlying Request Format

EndpointR handles the JSON formatting for you, but for debugging purposes, it is helpful to know what the API expects. Each line in the batch file is a JSON object containing a custom_id and the request body.


```{json}         
{
    "custom_id": "doc_1",
    "method": "POST",
    "url": "/v1/embeddings",
    "body": {
        "input": "The quick brown fox...",
        "model": "text-embedding-3-small",
        "encoding_format": "float"
    }
}
```
