---
title: "Synchronous vs Asynchronous APIs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synchronous vs Asynchronous APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EndpointR)
```

# Introduction

Most of EndpointR's integrations are with synchronous APIs such as [Completions](https://platform.openai.com/docs/api-reference/completions) by OpenAI, Hugging Face's [Inference Endpoints](https://huggingface.co/docs/inference-endpoints/en/index), and Messages by [Anthropic](https://platform.claude.com/docs/en/api/messages). When using these APIs, we send a request, wait a second or two and receive a response.

However, most Generative AI providers also offer lower-cost, asynchronous APIs. The providers usually offer a guarantee of the results within a time frame, and an estimate of the average time to return the results. For example, they may guarantee results within 24 hours, but expect them within 1-3 hours.

# When to choose Synchronous vs Asynchronous

For us as consumers, the decision is a trade-off between time and money. If we are serving Generative AI to other consumers, e.g. in an application, we will usually favour a Synchronous API because users expect instant results. Alternatively, if we are running analyses over large datasets, or repeated batch-inference, we can usually afford to wait longer for the results so we may favour the Asynchronous APIs.

Synchronous APIs are also very useful when we are still in the experimental step of an analysis and need quick feedback- i.e. when we're testing our prompts, and developing our [schemas](https://json-schema.org/) for [Structured Outputs](https://platform.openai.com/docs/guides/structured-outputs), we don't want to wait 24 hours to check whether our schema works. Instead, we want to send a request and receive a response within seconds. That way we can iteratively fix/develop our schemas/prompts and get to a better outcome quicker.
