---
title: "openai_batch_api"
format: html
---

So... we could actually re-use some of the logic from oai_embed.R / oai_classify.R. There are some small differences, e.g. we feed a stub of the endpoint URL to each request, and then we upload the batch of inputs as a file directly to the files API, then create the batch.

```{r}
embed_req <- oai_build_embedding_request("xx", dimensions = 324)

body <- embed_req$body$data

row <- list(
  custom_id = "xx",
  method = "POST",
  url = "/v1/embeddings",
  body = embed_req$body$data
)
```

Initial thought was to just stream_out, but stream_out expects a data frame as input. Which we can use further down the line, but not here.

```{r}
jsonlite::stream_out(row, con = stdout()) 

tib_w_row <- tibble::tibble(rows = list(row))

jsonlite::stream_out(tib_w_row, con = file("test_dir/jsonl_outputs/batch_api_test.jsonl"))

read_in <- jsonlite::stream_in(con = file("test_dir/jsonl_outputs/batch_api_test.jsonl")) |> jsonlite::toJSON()
read_in["rows"]

tibble::tibble(x = 1:10^5) |> 
  chunk_dataframe(chunk_size = 80000)
```

So instead, we want to take the row, convert it to JSON with `auto_unbox = TRUE` and then writeLines to a `.jsonl` file. Recall that `auto_unbox` just stops each k:v pair's value being treated as a list

```{r}
# stream in./ out not the right way
jsonlite::toJSON(row, auto_unbox = TRUE) |>
  writeLines("test_dir/jsonl_outputs/batch_api_test_write_lines.jsonl")
  

x <- readLines("test_dir/jsonl_outputs/batch_api_test_write_lines.jsonl")
jsonlite::toJSON(x, auto_unbox = TRUE)
```

We said we could re-use some of the logic, but looking at it we don't really benefit from using httr2 for each request - it's unnecessary overhead. So we just create lists for now. We may use httr2 for the actual batch request (but maybe not!)

```{r}
single_batch_row <- oai_batch_build_embed_req("hello", "1")

list_rows <- purrr::map(1:10, \(x) oai_batch_build_embed_req("hello", x))
```

Then we can write them to a file as follows, and send it to OpenAI as a batch job.

```{r}
writeLines(
  unlist(list_rows), 
  "test_dir/jsonl_outputs/write_ten_lines.jsonl")


readLines(
  "test_dir/jsonl_outputs/write_ten_lines.jsonl",
  n = 2
)
```

```{r}
test_df <- tibble::tibble(
  x = letters,
  y = 1:length(letters)
)

test_df |>
  mutate(
    reqs = map2_chr(x, y, \(text, id) {
      oai_batch_build_embed_req(
        text,
        id,
        dimensions = 324
      )
    })
  )

xx <- test_df |>
  oai_batch_prepare_embeddings(
    x,
    y
  )

oai_batch_file_upload(
  xx
)

batch_job_data <- oai_batch_file_list()
temp_id <- batch_job_data$data[[1]]$id

oai_batch_file_delete(temp_id)

oai_batch_file_list()
```
