---
title: "openai_batch_api"
format: html
---

So... we could actually re-use some of the logic from oai_embed.R / oai_classify.R. There are some small differences, e.g. we feed a stub of the endpoint URL to each request, and then we upload the batch of inputs as a file directly to the files API, then create the batch.

```{r}
embed_req <- oai_build_embedding_request("xx", dimensions = 324)

body <- embed_req$body$data

row <- list(
  custom_id = "xx",
  method = "POST",
  url = "/v1/embeddings",
  body = embed_req$body$data
)
```

Initial thought was to just stream_out, but stream_out expects a data frame as input. Which we can use further down the line, but not here.

```{r}
jsonlite::stream_out(row, con = stdout()) 

tib_w_row <- tibble::tibble(rows = list(row))

jsonlite::stream_out(tib_w_row, con = file("test_dir/jsonl_outputs/batch_api_test.jsonl"))

read_in <- jsonlite::stream_in(con = file("test_dir/jsonl_outputs/batch_api_test.jsonl")) |> jsonlite::toJSON()
read_in["rows"]

tibble::tibble(x = 1:10^5) |> 
  chunk_dataframe(chunk_size = 80000)
```

So instead, we want to take the row, convert it to JSON with `auto_unbox = TRUE` and then writeLines to a `.jsonl` file. Recall that `auto_unbox` just stops each k:v pair's value being treated as a list

```{r}
# stream in./ out not the right way
jsonlite::toJSON(row, auto_unbox = TRUE) |>
  writeLines("test_dir/jsonl_outputs/batch_api_test_write_lines.jsonl")
  

x <- readLines("test_dir/jsonl_outputs/batch_api_test_write_lines.jsonl")
jsonlite::toJSON(x, auto_unbox = TRUE)
```

We said we could re-use some of the logic, but looking at it we don't really benefit from using httr2 for each request - it's unnecessary overhead. So we just create lists for now. We may use httr2 for the actual batch request (but maybe not!)

```{r}
single_batch_row <- oai_batch_build_embed_req("hello", "1")

list_rows <- purrr::map(1:10, \(x) oai_batch_build_embed_req("hello", x))
```

Then we can write them to a file as follows, and send it to OpenAI as a batch job.

```{r}
writeLines(
  unlist(list_rows), 
  "test_dir/jsonl_outputs/write_ten_lines.jsonl")


readLines(
  "test_dir/jsonl_outputs/write_ten_lines.jsonl",
  n = 2
)
```

```{r}
test_df <- tibble::tibble(
  x = letters,
  y = 1:length(letters)
)

test_df |>
  mutate(
    reqs = map2_chr(x, y, \(text, id) {
      oai_batch_build_embed_req(
        text,
        id,
        dimensions = 324
      )
    })
  )

xx <- test_df |>
  oai_batch_prepare_embeddings(
    x,
    y
  )

oai_batch_upload(
  xx
)

batch_job_data <- oai_batch_file_list()
temp_id <- batch_job_data$data[[1]]$id

oai_batch_file_delete(temp_id)

oai_batch_file_list()
```

# Testing

## Embeddings

```{r}
uploaded_files <- oai_file_list(
  purpose = "batch"
)
assertthat::validate_that(length(uploaded_files$data) == 0)
```

First batch failed due to ID beingan integer, so ID has to be a string...

```{r}
embedding_rows <- test_df |>
  oai_batch_prepare_embeddings(
    x,
    y
  )

embedding_file <- oai_batch_upload(embedding_rows)

embedding_batch <- oai_batch_start(embedding_file$id,
                 endpoint = "/v1/embeddings")

batch_jobs <- oai_batch_list()
batch_jobs
oai_batch_status(embedding_batch$id)
```

And then this time, so we need to fix how the tmpfile is created and prepend with batch:

```         
$error$message
[1] "Invalid 'batch_id': 'file-K6vaHgwcJsE5z1MMFvVMix'. Expected an ID that begins with 'batch'."
```

Once we've made sure the ID is a string, then we create the file, upload it, start the batch, check the status, download it.

I would say that this still feels quite janky. There's a few different file ids going on, and then we need to handle any errors separately it seems?

```{r}
embedding_rows_id_string <- test_df |>
  oai_batch_prepare_embeddings(
    x,
    y
  )

embedding_file_id_string <- oai_batch_upload(embedding_rows_id_string)

embedding_batch_id_string <- oai_batch_start(embedding_file_id_string$id,
                 endpoint = "/v1/embeddings")

embedding_batch_metadata <- oai_batch_status(
  embedding_batch_id_string$id # "batch_6960e0b48bf481909751c76756ac9fec"
)

output_file_contents <- oai_file_content(
  embedding_batch_metadata$output_file_id
)


oai_batch_parse_embeddings(output_file_contents, original_df = NULL)
```

## Low dimensions

Looks good, 327 columns = 324 + 3 (id, .error, .error_msg)

```{r}
ld_embed_rows <- test_df |>
  oai_batch_prepare_embeddings(
    x,
    y,
    dimensions = 324
  )

ld_file <- oai_batch_upload(
  jsonl_rows = ld_embed_rows,
  purpose = "batch"
)

ld_batch_job <- oai_batch_start(ld_file$id, endpoint = "/v1/embeddings")

oai_batch_status(ld_batch_job$id)[["status"]]
ld_results <- oai_file_content(oai_batch_status(ld_batch_job$id)[["output_file_id"]])

oai_batch_parse_embeddings(ld_results)
  
```

Delete/clean up

We can delete the input/output files, but it doesn't seem like we can actually delete batches.

```{r}
oai_batch_list()[["data"]] |> 
  map( pluck("id")
) |>  unlist()

oai_batch_status("batch_6960ed03eea8819080aaa69a8982de66")
oai_file_delete("file-4f8STaon74XE5yP6M7mWmH")
oai_file_delete("file-FteMgeWV4mK85kdcntGU29")

oai_batch_status("batch_6960dfe7d8ac8190acf682a59e844b71")
oai_file_delete("file-HRFKx63PimYZYqsraQHR6T")
oai_file_delete("file-HRFKx63PimYZYqsraQHR6T")
oai_batch_status("batch_6960e0b48bf481909751c76756ac9fec")
oai_file_delete('file-TcV6dGkFEsrzGTNxkPKYCb') # output file
oai_file_delete('file-K6vaHgwcJsE5z1MMFvVMix') # input file

```

## Completions

testing the funcs for Completions, we need to get an input, make a file, start the batch, checl the status, retrieve the content
```{r}
completions_req <- oai_batch_build_completions_req(
  input = "Tell me a joke about my country, the United Kingdom",
  id = "id_1"
)

completions_file <- oai_batch_upload(
  completions_req
)
```

Do need to remember to fill in endpoint = "/v1/chat/completions" here instead of default arg for embeddings
```{r}
completions_batch <- oai_batch_start(
  completions_file$id,
  endpoint = "/v1/chat/completions"
)

completions_status <- oai_batch_status(
  completions_batch$id
)

completions_status$status
completions_status$output_file_id # file-Q8TaFRoCYGHRZJKiYQKhx9

output <- oai_file_content(completions_status$output_file_id)

oai_batch_parse_completions(output) |> 
  purrr::pluck("content", 1)
```

### With Schema

```{r}
joke_schema <- create_json_schema(
  name = "joke_schema",
  description = "A set up and a punchline",
  schema = schema_object(
    setup = schema_string("The set up for the joke"),
    punchline = schema_string("The punchline of the joke, make it pop"),
    required = c("setup", "punchline")
  )
)

joke_schema

completions_req_w_schema <- oai_batch_build_completions_req(
  input = "Tell me a joke about my country, the United Kingdom",
  id = "id_1",
  schema = joke_schema,
  temperature = 1
)

.file <- oai_batch_upload(
  completions_req_w_schema
)

.batch <- oai_batch_start(.file$id, endpoint = "/v1/chat/completions")

oai_batch_status(.batch$id)[["status"]]

oai_batch_status(.batch$id)[["output_file_id"]] |> 
  oai_file_content()
oai_batch_status(.batch$id)[[""]]
```

To deal with a batch output that had strutured outputs, we need to first parse the batch, and then use the same approach we use elsewhere in Endpoint  - may work on this as it still feels quite clunky
```{r}
.content <- oai_batch_status(.batch$id)[["output_file_id"]] |> 
  oai_file_content()

parsed_batch <- oai_batch_parse_completions(.content) # custom_id, content, .error, .error_msg

parsed_batch |>
    dplyr::mutate(
      parsed = purrr::map(content, 
      \(x) safely_from_json(x))
    ) |>
    tidyr::unnest_wider(parsed)
```


# Files API

Write the generic oai_file_upload func and then call that it in oai_batch_upload
Rename oai_batch_file_upload --> oai_batch_upload(?)

```{r}
tmp <- tempfile(fileext = ".jsonl")
writeLines("Hello!", tmp)
readLines(tmp)
file.path(tmp)

test_upload <- oai_file_upload(
  file = tmp,
  purpose = "user_data"
) # file must be a file object

file(tmp)
```

```{r}

```